A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications

### 2014-05-26 Paper/paper.tex
New format, thanks Tim! Issues with figures though....

% Two Column Format
\documentclass[11pt]{article}
%this allows us to specify sections to be single or multi column so that things like title page and table of contents are single column
\usepackage{multicol, caption}
\usepackage{verbatim}
\usepackage{listings}
\usepackage{color}
\usepackage[disable]{todonotes}

@@

\usepackage{setspace}
\usepackage{url}

\usepackage{graphicx}

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{letterpaper}

\setcounter{secnumdepth}{5}

@@

\newenvironment{Figure}
  {\par\medskip\noindent\minipage{\linewidth}}
  {\endminipage\par\medskip}

@@

\title{\vfill A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications} %\vfill gives us the black space at the top of the page
\author{
By Taggart Ashby \vspace{10pt} \\
Computer Science Department, California Polytechnic State University, San Luis Obispo
}

@@

\vfill  %in combination with \newpage this forces the abstract to the bottom of the page

@@

\thispagestyle{empty} %remove page number from title page
\newpage

%end the 1 column format

%start 2 column format
% \begin{multicols}{2}
%Start numbering first page of content as page 1
\setcounter{page}{1}

%%%%%%%%%%%%%%%%%%%%
%%% Introduction/Motivation
%%% More detailed, but still general, description of problem and why we want to solve it
%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Since the advent of the internet, the web has undergone an impressive evolution from plain-text webpages to the highly stylized and functional pages that we see today. In the last four to six years there has been an explosion of new web technologies that make applications more functional: HTML5, CSS3, WebGL, Touch APIs, Geo-location, and a host of others. \cite{EvolutionOfWeb} In that same time period, the number of global internet users has grown to somewhere around two and a half billion. \cite{EvolutionOfWeb} The combination of these new technologies and the ever-increasing usage of the internet has led to a significant growth in the number and complexity of web applications.

@@

\subsection{What Has Been Done Already?}
Web developers are not clueless, they understand that there is a need for testing of web applications. Unfortunately, web development is still a relatively new domain and developers have not come together to work on the problem of testing but rather have come up with their own proprietary solutions in isolation or merely on a per-project basis. Some of these tools have gained traction and become larger, more general, tools.

@@

\subsubsection{Why It Is Insufficient}
% - No standard
The problem with proprietary tools is that many of the tools attempt to remedy the same problems. A developer has to decide between a number of similar tools that all solve similar problems. Having to wade through a sea of similar tools is time-consuming, which can be the death of a cutting-edge application. The more time an application is off the market, the more time competitors have to develop their own versions. Yet another problem is that deciding between similar tools can be discouraging to the developer that just wants to write their application and test it quickly.

@@

\subsection{Problems Still Faced/Why It's So Hard}
% - A lot of products to test against (browsers)
Testing web applications is difficult. The web offers an impressive set of features, but with those features come some draw-backs. Two of the most frustrating drawbacks for developers are the number of platforms and the web's asynchronous nature.

@@

\subsection{Contributions of This Thesis}
%- Framework
The goal of this thesis is to bring together tools that are already available for testing and combine them into a single framework that offers web application developers a toolkit for their testing needs. There are plenty of strong tools available and we hope to gather a subset of those tools to cover as much of the testing spectrum as possible.

@@

%%%%%%%%%%%%%%%%%%
%%% Background 
%%% Information needed for reader to know what’s going on
%%% 	Server/Client Communication
%%% 	Cloud services
%%% 	Appropriate terms and definitions
%%%%%%%%%%%%%%%%%%
\section{Background}
In an effort to make the rest of the paper more understandable and to define the terms used in the remainder of the paper, we will outline the various concepts and technologies surrounding the web.
If one is familiar with HTML5, JavaScript, and Testing concepts and statistics, it will likely be safe to skip this background section.

\subsection{Web Application Strengths}
Web applications are great for both developers and consumers. They have a number of characteristics that make them more desirable than shrink-wrap software. These attributes include versioning, availability, platform independence, and the ability to rapidly prototype.

\subsubsection{Versioning}
Web applications can be updated at any time with instant roll-out and feedback. It requires absolutely no user interaction because the web page always serves the latest resources. In comparison to something like an operating system that requires users to manually update, this feature leads to applications that can address problems quickly, roll-out security updates instantly, and incorporate consumer feedback much more quickly than any other type of software. Another perk is that users are not given a choice. This can lead to occasional customer outrage, but more often than not it's a powerful gain for both feature improvement and application security.

\subsubsection{Availability}
Web application are available anywhere there's internet. There is no need to install anything on a given device. Due to the ever-increasing popularity of mobile devices, most applications have both mobile and full versions of their software and so the application is ``with'' the user all the time. This instant and continuous availability means that users only have to know how to use a web browser in order to access the application.

\subsubsection{Multi-platform}
Very similar to the availability strength, web applications are platform independent. Whether the user is on their phone, tablet, laptop, or PC, the application only requires a web browser with certain functionality and in this day and age all devices come equipped with browsers that should have no problems. This also means that the developer need not, generally, worry about hardware configurations, conflicting software, and other problems that can arise in desktop applications.

\subsubsection{Rapid Prototyping}
Web applications are fantastic for rapid prototyping because of all of the strengths discussed above. An idea can be prototyped and posted on a website quickly and rapid iteration can occur without the user even necessarily noticing. There are a number of web tools that will record where users are clicking, what features they're using, how long they're spending on a given page or step of a process, and a number of other metrics that can make iterations more directed and effective.

\subsection{Web Features}
Here we'll discuss some of the prevalent technologies that recently appeared in web applications and development. By no means is this an exhaustive list, just a sampling of the more prominent ones.

\subsubsection{HTML5}
The latest revision of HTML is HTML5 which first debuted on Firefox back in 2009. \cite{EvolutionOfWeb} HTML5 has not been officially recommended by the W3C (World Wide Web Consortium) who is the official keeper and recommender of web standards, however, there is a plan for final recommendation this year, 2014, called Plan 2014. \cite{Plan2014} What this means is that in 2014 the W3C will release a stable HTML5 Recommendation which will make HTML5 an official standard and help universalize all the separate implementations.

@@

%\todo{THIS NEEDS TO MOVE OR CHANGE SUBSECTION TO `WEB APPLICATIONS'}
\paragraph{Node.JS}
As a quick aside, it will be useful to have information about Node.JS as it is used in the sample application and the evaluation application. Node.JS, often called simply Node, is a JavaScript API for creating and running servers. ``Node's goal is to provide an easy way to build scalable network programs.'' \cite{Node} Node was created in 2009 and is sponsored by Joyent. Node does not use thread-based networking, instead opting for a single-threaded event loop and non-blocking I/O. Node allows developers to create and control web servers without the need for external software, such as Apache, commonly found in other sites.

@@

\todo{TIE THIS CLIENT/SERVER STUFF into Web Applications, be specific, not generic}
\subsection{Client/Server Architecture}
Web applications have a user-facing client-side and a ``rear-facing'' \todo{IS THIS THE RIGHT TERM...} server-side.

@@

\subsubsection{What is the Client-side Portion?}
First, we'll address what a client is. The simple answer is: The User. The User's phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service being using to communicate with the web is the client.

@@

\subsubsection{What is the Server-side Portion?}
Once again, we'll outline a server and then delve into server-side programming. A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets, HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and a general number-crunching.

@@

\subsubsection{Client and Server Interactions}
The client has limited access to resources on any given web application. This access is restricted by the server. Without such restrictions in place, someone could access user databases, credit card information, order reports, etc. In addition, there would be an overwhelming amount of content for a client to parse through, even if they have no malicious intent.

@@

\subsection{What is Testing?}
IEEE defines software testing as ``the dynamic verification of the behavior of a program on a finite set of test cases, suitably selected from the usually infinite executions domain, against the expected behavior.'' \cite{TestingDefinition} Put a bit more simply, software testing is the process of making sure the program does what it is supposed to.

@@

\subsubsection{Unit Testing}
Unit testing is the testing of specific functions and functionality, or ``test units.'' Test units are ``A set of one or more computer program modules together with associated control data, [...] usage procedures, and operating procedures that satisfy the following conditions: 1) All modules are from a single computer program. 2) At least one of the new or changed modules in the set has not completed the unit test. 3) The set of modules together with its associated data and procedures are the sole object of a testing process.'' \cite{UnitTestDefinition} Generally this means the testing of individual functions but can extend to a set of functions that produce a single module of functionality. What sets Unit Testing apart from other types of testing is that tests focus on one portion of the code or system rather than interactions between modules or the entire system.

@@

\subsubsection{Continuous Integration}
The original definition of Continuous Integration involves the idea of developers committing from their  branches to a master branch daily so that changes are continuously integrated in an effort to make sure all branches can, in fact, be integrated. It has since evolved into a more general idea of a tool that runs commands after each commit to a repository. These commands generally encompass running unit tests, integration tests, user interface tests, and also often publishing and disseminating notifications to interested parties. This is invaluable feedback because the developer can see if they're making progress in the right direction as they write code, as opposed to a situation in which the developer finishes coding and finds out they've made a bad assumption or decision between this commit and the last.

@@

\subsubsection{Acceptance Testing}
Acceptance testing is testing to confirm that the product meets a certain specification. Often this is making sure that the software meets an SRS (Software Requirements Specification) or a customer's outlined expectations. Often this testing is performed by the user of the software or customer that ordered the software, rather than the developers.

@@

\subsubsection{Testing Statistics}
It is widely accepted that the earlier a developer detects a defect, the easier it is to fix that defect. A famous book, ``Code Complete'' by Steve McConnell \cite{DefectPic}, includes the following figure, showing this statistic on a rough graph.

\begin{Figure}
	\centering
	\includegraphics[width=0.75\linewidth]{defectcost.jpg} 
	\captionof{figure}{\cite{DefectPic}}
\end{Figure}

@@

% - CONSIDER REMOVING ALTOGETHER
% \subsubsection{Why Testing is Important}
% We're consistently surprised by how few people take testing seriously when it comes to non-trivial applications. Many developers will leave testing to their users, i.e., wait for users to have problems and report them before developers will fix them. This is an issue for a number of reasons, a few of which are: frustrated customers are unlikely to continue using a buggy product and are dissuaded from buying future products from the same developer, certain bugs can be security loopholes that may allow users access to information they shouldn't have, in the worst cases a bug may cause damage to a person or property and the developers would be on the hook.

%%%%%%%%%%%%%%%%%%%%%%
%%% Sample App 
%%%   Talk about test app
%%%%%%%%%%%%%%%%%%%%%%
\section{Sample App and Testing Frameworks}

\subsection{Sample ToDo App}
In order to evaluate the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our evaluation application. After some searching, we decided upon a Todo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} Here is a brief preview of the app:

\begin{Figure}
  \centering
  \includegraphics[width=0.75\linewidth]{todo_home.png}
  \captionof{figure}{ToDo app home page, simple three-button naviagation}
\end{Figure}

\begin{Figure}
  \centering
  \includegraphics[width=0.75\linewidth]{todo_some_tasks.png}
  \captionof{figure}{ToDo app task page}
\end{Figure}

\begin{Figure}
  \centering
  \includegraphics[width=0.75\linewidth]{todo_completed.png}
  \captionof{figure}{ToDo app completed tasks page}
\end{Figure}

@@

%%%%%%%%%%%%%%%%%%%%%%
%%% Testing Frameworks
%%%   Talk about the various frameworks
%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing Frameworks}

% NOT SURE I NEED THIS SECTION AT ALL
% \subsection{Assumptions and Requirements}
% In order to more thoroughly vet the many testing frameworks we came across, we had a number of assumptions about the projects we'd be looking at and requirements for those framework. 
%- Node.js server because it's JS, so it's testable.
%- Something to do unit testing
%- Something to do continuous integration

\subsubsection{The Average Web Application}
In an effort to create a good workflow and collection of tools, we have chosen to focus on web applications that uses Node.js, JavaScript, Cloud Services, and a backing Database when searching for our tools. This choice was made because of the popularity of that kind of set up, as well as the fact that our evaluation application was set up in that way.

@@

\subsubsection{Tool Requirements}
\todo{Move this section elsewhere}
Our requirements for the selection of testing tools was relatively simple. In addition to meeting the requirements of our ``average web application'', the only other requirement was that the tool was free. What ``free'' means is that the developer would have access to at least the minimum functionality needed to use the tool on \emph{one} application that is \emph{open-source}. What we quickly found is that web testing tools love open-source projects but are far less willing to provide free functionality to private or closed-source projects.

@@

\subsection{Unit Testing Frameworks}
The following frameworks are reviewed in chronological order of our evaluation. As such, comparisons will generally only be drawn between a framework and its predecessors in paper order. After all of the frameworks are outlined and evaluated, a longer section regarding the final decision will be made where comparisons will be drawn between all of them.

@@

\subsubsection{Jasmine \cite{Jasmine}}
The first of the unit testing frameworks that we looked at was Jasmine, a Behavior Driven Development (BDD) testing framework. BDD is based on Test Driven Development and adds some simplifications and patterns that attempt to bring both developers and businessmen into the software testing process. The idea behind Behavior Driven Development is that input from non-technical stakeholders as well as technical stakeholders can come together so that everyone understands what the project should do.

@@

\subsubsection{Mocha \cite{Mocha}}
Mocha is another BDD testing framework that is almost identical to Jasmine. Unlike Jasmine, however, Mocha is not completely self-contained and requires an additional assertion library of one's choosing. We will outline our choice of Chai.js below.

@@

\paragraph{Assertion Library of Choice: Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of \lstinline{expect(myVar).to.equal(10);} or \lstinline{expect(myArr).to.have.length(10);}. The Should API has assertions of the form \lstinline{myVar.should.be.a(\'number\');} or \lstinline{myArr.should.have.length(10);}. Finally, the Assert API has the more tried and true \lstinline{assert.equal(myVar, 10);} and \lstinline{assert.lengthOf(myArr, 10);}. The various APIs combined with Mocha's flexibility led to a nice developer experience that can appeal to a variety of different developers.

\subsubsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the \lstinline{it} functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

\subsubsection{Intern\cite{InternIO}}
Intern is a hybrid unit testing and browser testing framework that aims to be an all-in-one testing framework. Intern has unit testing, browser testing, and code coverage information all built into the same tool. Intern runs on its own server which give it access to the internals of the project, but caused problems when attempting to interact with a separate server running the Node.js application. We were unable to get the browser testing portion of Intern to work because of the inability to set up cross-server communication easily and properly.

@@

\subsection{Browser and UI Testing}
Web applications, for the most part, require user input in order to be useful. The way the user interacts is through the outward facing web pages. Testing of a web application would be incomplete without testings the interactions and elements of these web pages. It's important that the inner workings are running the right code and producing the right answer, but it's almost just as important that the parameters get passed correctly and from the right places on the web pages. Also, it is important that navigation work as expected or a user will become frustrated and lose interest in the application.

@@

\subsubsection{Selenium\cite{Selenium}}
Selenium is an automated browser input runner. It is one of the only automated browser runners and certainly the most prominent. A developer can write code that will cause Selenium to drive a browser's input and assert expectations. On its own Selenium has very minimal testing usage. It lacks a test reporting mechanism and only offers output in two forms: 1) No output at all, a browser window will open, input will be generated, and the window will close, or 2) the browser will exit with a cryptic error printed to the console. It is up to the developer to include print statements as to what is happening at any given time, what test is being run, and what the results are.

@@

\subsubsection{Nightwatchjs\cite{NightwatchJS}}
Nightwatch is a JavaScript testing library that wraps Selenium in an effort to make browser testing more concise and straightforward. Nightwatch uses Node.js' assertion library paired with Selenium's WebDriver to both run browser automation and also report test results.

@@

\subsection{Testing Frameworks}

>>>>

% Two Column Format
\documentclass[12pt]{ucthesis}
%this allows us to specify sections to be single or multi column so that things like title page and table of contents are single column
\usepackage{listings}
\usepackage[disable]{todonotes}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[letterpaper]{geometry}
\usepackage[overload]{textcase}
\usepackage{verbatim}

\usepackage{titlesec}

@@

\titleformat{\chapter}[display]% NEW
    {\normalfont\begin{center}}{\chaptertitlename\ \thechapter}{12pt}{}% NEW
\titlespacing*{\chapter}{0pt}{30pt}{20pt}% NEW

%\titleformat{\chapter}[block]{first}{label}{12pt}

\titleformat{\chapter}{}{\thesection}{1em}{}
\titleformat{\section}{}{\thesubsection}{1em}{}
\titleformat{\subsection}{}{\thesubsubsection}{1em}{}
\titleformat{\subsubsection}{}{\theparagraph}{1em}{}

% \setcounter{secnumdepth}{5}
\setlength{\parindent}{0.25in} \setlength{\parskip}{6pt}

\geometry{verbose,nohead,tmargin=1.25in,bmargin=1in,lmargin=1.5in,rmargin=1.3in}

\setcounter{tocdepth}{2}

@@

% Different font in captions (single-spaced, bold) ------------
\newcommand{\captionfonts
\captionfonts}{\small\bf\ssp}

\makeatletter  % Allow the use of @ in command names
\long\def\@makecaption#1#2{%
  \vskip\abovecaptionskip
  \sbox\@tempboxa{{\captionfonts
\captionfonts #1: #2}}%
  \ifdim \wd\@tempboxa >\hsize
    {\captionfonts
\captionfonts #1: #2\par}
  \else
    \hbox to\hsize{\hfil\box\@tempboxa\hfil}%
  \fi
  \vskip\belowcaptionskip}
\makeatother   % Cancel the effect of \makeatletter
% ---------------------------------------

@@

\title{A Survey of Web Testing Frameworks and Workflow for Testing JavaScript Applications} %\vfill gives us the black space at the top of the page
\author{Taggart Ashby}
\degreemonth{June} \degreeyear{2014} \degree{Master of Science}
\defensemonth{June} \defenseyear{2014}
\numberofmembers{3} \chair{Michael Haungs, Ph.D.} \othermemberA{Franz Kurfess, Ph.D.} \othermemberB{David Janzen, Ph.D.} \field{Computer Science} \campus{San Luis Obispo}
\copyrightyears{seven}

@@

\begin{frontmatter}

\copyrightpage

\committeemembershippage

@@

\tableofcontents

\end{frontmatter}

\pagestyle{plain}

\renewcommand{\baselinestretch}{1.66}

%%%%%%%%%%%%%%%%%%%%
%%% Introduction/Motivation
%%% More detailed, but still general, description of problem and why we want to solve it
%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
Since the advent of the internet, the web has undergone an impressive evolution from plain-text webpages to the highly stylized and functional pages that we see today. In the last four to six years there has been an explosion of new web technologies that make applications more functional: HTML5, CSS3, WebGL, Touch APIs, Geo-location, and a host of others. \cite{EvolutionOfWeb} In that same time period, the number of global internet users has grown to somewhere around two and a half billion. \cite{EvolutionOfWeb} The combination of these new technologies and the ever-increasing usage of the internet has led to a significant growth in the number and complexity of web applications.

@@

\section{What Has Been Done Already?}
Web developers are not clueless, they understand that there is a need for testing of web applications. Unfortunately, web development is still a relatively new domain and developers have not come together to work on the problem of testing but rather have come up with their own proprietary solutions in isolation or merely on a per-project basis. Some of these tools have gained traction and become larger, more general, tools.

@@

\subsection{Why It Is Insufficient}
% - No standard
The problem with proprietary tools is that many of the tools attempt to remedy the same problems. A developer has to decide between a number of similar tools that all solve similar problems. Having to wade through a sea of similar tools is time-consuming, which can be the death of a cutting-edge application. The more time an application is off the market, the more time competitors have to develop their own versions. Yet another problem is that deciding between similar tools can be discouraging to the developer that just wants to write their application and test it quickly.

@@

\section{Problems Still Faced/Why It's So Hard}
% - A lot of products to test against (browsers)
Testing web applications is difficult. The web offers an impressive set of features, but with those features come some draw-backs. Two of the most frustrating drawbacks for developers are the number of platforms and the web's asynchronous nature.

@@

\section{Contributions of This Thesis}
%- Framework
The goal of this thesis is to bring together tools that are already available for testing and combine them into a single framework that offers web application developers a toolkit for their testing needs. There are plenty of strong tools available and we hope to gather a subset of those tools to cover as much of the testing spectrum as possible.

@@

%%%%%%%%%%%%%%%%%%
%%% Background 
%%% Information needed for reader to know what’s going on
%%% 	Server/Client Communication
%%% 	Cloud services
%%% 	Appropriate terms and definitions
%%%%%%%%%%%%%%%%%%
\chapter{Background}
In an effort to make the rest of the paper more understandable and to define the terms used in the remainder of the paper, we will outline the various concepts and technologies surrounding the web.
If one is familiar with HTML5, JavaScript, and Testing concepts and statistics, it will likely be safe to skip this background section.

\section{Web Application Strengths}
Web applications are great for both developers and consumers. They have a number of characteristics that make them more desirable than shrink-wrap software. These attributes include versioning, availability, platform independence, and the ability to rapidly prototype.

\subsection{Versioning}
Web applications can be updated at any time with instant roll-out and feedback. It requires absolutely no user interaction because the web page always serves the latest resources. In comparison to something like an operating system that requires users to manually update, this feature leads to applications that can address problems quickly, roll-out security updates instantly, and incorporate consumer feedback much more quickly than any other type of software. Another perk is that users are not given a choice. This can lead to occasional customer outrage, but more often than not it's a powerful gain for both feature improvement and application security.

\subsection{Availability}
Web application are available anywhere there's internet. There is no need to install anything on a given device. Due to the ever-increasing popularity of mobile devices, most applications have both mobile and full versions of their software and so the application is ``with'' the user all the time. This instant and continuous availability means that users only have to know how to use a web browser in order to access the application.

\subsection{Multi-platform}
Very similar to the availability strength, web applications are platform independent. Whether the user is on their phone, tablet, laptop, or PC, the application only requires a web browser with certain functionality and in this day and age all devices come equipped with browsers that should have no problems. This also means that the developer need not, generally, worry about hardware configurations, conflicting software, and other problems that can arise in desktop applications.

\subsection{Rapid Prototyping}
Web applications are fantastic for rapid prototyping because of all of the strengths discussed above. An idea can be prototyped and posted on a website quickly and rapid iteration can occur without the user even necessarily noticing. There are a number of web tools that will record where users are clicking, what features they're using, how long they're spending on a given page or step of a process, and a number of other metrics that can make iterations more directed and effective.

\section{Web Features}
Here we'll discuss some of the prevalent technologies that recently appeared in web applications and development. By no means is this an exhaustive list, just a sampling of the more prominent ones.

\subsection{HTML5}
The latest revision of HTML is HTML5 which first debuted on Firefox back in 2009. \cite{EvolutionOfWeb} HTML5 has not been officially recommended by the W3C (World Wide Web Consortium) who is the official keeper and recommender of web standards, however, there is a plan for final recommendation this year, 2014, called Plan 2014. \cite{Plan2014} What this means is that in 2014 the W3C will release a stable HTML5 Recommendation which will make HTML5 an official standard and help universalize all the separate implementations.

@@

%\todo{THIS NEEDS TO MOVE OR CHANGE SUBSECTION TO `WEB APPLICATIONS'}
\subsubsection{Node.JS}
As a quick aside, it will be useful to have information about Node.JS as it is used in the sample application and the evaluation application. Node.JS, often called simply Node, is a JavaScript API for creating and running servers. ``Node's goal is to provide an easy way to build scalable network programs.'' \cite{Node} Node was created in 2009 and is sponsored by Joyent. Node does not use thread-based networking, instead opting for a single-threaded event loop and non-blocking I/O. Node allows developers to create and control web servers without the need for external software, such as Apache, commonly found in other sites.

@@

\todo{TIE THIS CLIENT/SERVER STUFF into Web Applications, be specific, not generic}
\section{Client/Server Architecture}
Web applications have a user-facing client-side and a ``rear-facing'' \todo{IS THIS THE RIGHT TERM...} server-side.

@@

\subsection{What is the Client-side Portion?}
First, we'll address what a client is. The simple answer is: The User. The User's phone, computer, tablet, or other device communicating with a given website or application is the client. More specifically, the browser or service being using to communicate with the web is the client.

@@

\subsection{What is the Server-side Portion?}
Once again, we'll outline a server and then delve into server-side programming. A server is the entity, be it machine, person, or cloud service, that stores the information about a given web page or application. The server stores all assets, HTML pages, images, videos, databases, etc, and provides them upon request to the client. The server also handles all of the processing for interactivity, accounts, and a general number-crunching.

@@

\subsection{Client and Server Interactions}
The client has limited access to resources on any given web application. This access is restricted by the server. Without such restrictions in place, someone could access user databases, credit card information, order reports, etc. In addition, there would be an overwhelming amount of content for a client to parse through, even if they have no malicious intent.

@@

\section{What is Testing?}
IEEE defines software testing as ``the dynamic verification of the behavior of a program on a finite set of test cases, suitably selected from the usually infinite executions domain, against the expected behavior.'' \cite{TestingDefinition} Put a bit more simply, software testing is the process of making sure the program does what it is supposed to.

@@

\subsection{Unit Testing}
Unit testing is the testing of specific functions and functionality, or ``test units.'' Test units are ``A set of one or more computer program modules together with associated control data, [...] usage procedures, and operating procedures that satisfy the following conditions: 1) All modules are from a single computer program. 2) At least one of the new or changed modules in the set has not completed the unit test. 3) The set of modules together with its associated data and procedures are the sole object of a testing process.'' \cite{UnitTestDefinition} Generally this means the testing of individual functions but can extend to a set of functions that produce a single module of functionality. What sets Unit Testing apart from other types of testing is that tests focus on one portion of the code or system rather than interactions between modules or the entire system.

@@

\subsection{Continuous Integration}
The original definition of Continuous Integration involves the idea of developers committing from their  branches to a master branch daily so that changes are continuously integrated in an effort to make sure all branches can, in fact, be integrated. It has since evolved into a more general idea of a tool that runs commands after each commit to a repository. These commands generally encompass running unit tests, integration tests, user interface tests, and also often publishing and disseminating notifications to interested parties. This is invaluable feedback because the developer can see if they're making progress in the right direction as they write code, as opposed to a situation in which the developer finishes coding and finds out they've made a bad assumption or decision between this commit and the last.

@@

\subsection{Acceptance Testing}
Acceptance testing is testing to confirm that the product meets a certain specification. Often this is making sure that the software meets an SRS (Software Requirements Specification) or a customer's outlined expectations. Often this testing is performed by the user of the software or customer that ordered the software, rather than the developers.

@@

\subsection{Testing Statistics}
It is widely accepted that the earlier a developer detects a defect, the easier it is to fix that defect. A famous book, ``Code Complete'' by Steve McConnell \cite{DefectPic}, includes the following figure, showing this statistic on a rough graph.

\begin{figure}
	\begin{center}
	\includegraphics[width=0.75\linewidth]{defectcost.jpg} 
	\captionfonts
\caption[figure]{Illustration relating phase bug found in with effort required to fix.}
\end{center}
\end{figure}

@@

% - CONSIDER REMOVING ALTOGETHER
% \subsection{Why Testing is Important}
% We're consistently surprised by how few people take testing seriously when it comes to non-trivial applications. Many developers will leave testing to their users, i.e., wait for users to have problems and report them before developers will fix them. This is an issue for a number of reasons, a few of which are: frustrated customers are unlikely to continue using a buggy product and are dissuaded from buying future products from the same developer, certain bugs can be security loopholes that may allow users access to information they shouldn't have, in the worst cases a bug may cause damage to a person or property and the developers would be on the hook.

%%%%%%%%%%%%%%%%%%%%%%
%%% Sample App 
%%%   Talk about test app
%%%%%%%%%%%%%%%%%%%%%%
\chapter{Sample App and Testing Frameworks}

\section{Sample ToDo App}
In order to evaluate the various frameworks that we found online, we needed a small application with some functionality, but without the inherent complexity that was going to come with our evaluation application. After some searching, we decided upon a Todo application made with Express.js, Node.js, and MongoDB. \cite{ToDoAppHomePage} Here is a brief preview of the app:

\begin{figure}
  \begin{center}
  \includegraphics[width=0.75\linewidth]{todo_home.png}
  \captionfonts
\caption[todo]{ToDo app home page, simple three-button naviagation}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.75\linewidth]{todo_some_tasks.png}
  \captionfonts
\caption[figure]{ToDo app task page}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.75\linewidth]{todo_completed.png}
  \captionfonts
\caption[figure]{ToDo app completed tasks page}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%
%%% Testing Frameworks
%%%   Talk about the various frameworks
%%%%%%%%%%%%%%%%%%%%%%
\section{Testing Frameworks}

% NOT SURE I NEED THIS SECTION AT ALL
% \section{Assumptions and Requirements}
% In order to more thoroughly vet the many testing frameworks we came across, we had a number of assumptions about the projects we'd be looking at and requirements for those framework. 
%- Node.js server because it's JS, so it's testable.
%- Something to do unit testing
%- Something to do continuous integration

\subsection{The Average Web Application}
In an effort to create a good workflow and collection of tools, we have chosen to focus on web applications that uses Node.js, JavaScript, Cloud Services, and a backing Database when searching for our tools. This choice was made because of the popularity of that kind of set up, as well as the fact that our evaluation application was set up in that way.

@@

\subsection{Tool Requirements}
\todo{Move this section elsewhere}
Our requirements for the selection of testing tools was relatively simple. In addition to meeting the requirements of our ``average web application'', the only other requirement was that the tool was free. What ``free'' means is that the developer would have access to at least the minimum functionality needed to use the tool on \emph{one} application that is \emph{open-source}. What we quickly found is that web testing tools love open-source projects but are far less willing to provide free functionality to private or closed-source projects.

@@

\section{Unit Testing Frameworks}
The following frameworks are reviewed in chronological order of our evaluation. As such, comparisons will generally only be drawn between a framework and its predecessors in paper order. After all of the frameworks are outlined and evaluated, a longer section regarding the final decision will be made where comparisons will be drawn between all of them.

@@

\subsection{Jasmine \cite{Jasmine}}
The first of the unit testing frameworks that we looked at was Jasmine, a Behavior Driven Development (BDD) testing framework. BDD is based on Test Driven Development and adds some simplifications and patterns that attempt to bring both developers and businessmen into the software testing process. The idea behind Behavior Driven Development is that input from non-technical stakeholders as well as technical stakeholders can come together so that everyone understands what the project should do.

@@

\subsection{Mocha \cite{Mocha}}
Mocha is another BDD testing framework that is almost identical to Jasmine. Unlike Jasmine, however, Mocha is not completely self-contained and requires an additional assertion library of one's choosing. We will outline our choice of Chai.js below.

@@

\subsubsection{Assertion Library of Choice: Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of \lstinline{expect(myVar).to.equal(10);} or \lstinline{expect(myArr).to.have.length(10);}. The Should API has assertions of the form \lstinline{myVar.should.be.a(\'number\');} or \lstinline{myArr.should.have.length(10);}. Finally, the Assert API has the more tried and true \lstinline{assert.equal(myVar, 10);} and \lstinline{assert.lengthOf(myArr, 10);}. The various APIs combined with Mocha's flexibility led to a nice developer experience that can appeal to a variety of different developers.

\subsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the \lstinline{it} functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{figure}
  \begin{center}
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionfonts
\caption[figure]{Each individual spec is further broken down into the assertions within it.}
\end{center}
\end{figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

\subsection{Intern\cite{InternIO}}
Intern is a hybrid unit testing and browser testing framework that aims to be an all-in-one testing framework. Intern has unit testing, browser testing, and code coverage information all built into the same tool. Intern runs on its own server which give it access to the internals of the project, but caused problems when attempting to interact with a separate server running the Node.js application. We were unable to get the browser testing portion of Intern to work because of the inability to set up cross-server communication easily and properly.

@@

\section{Browser and UI Testing}
Web applications, for the most part, require user input in order to be useful. The way the user interacts is through the outward facing web pages. Testing of a web application would be incomplete without testings the interactions and elements of these web pages. It's important that the inner workings are running the right code and producing the right answer, but it's almost just as important that the parameters get passed correctly and from the right places on the web pages. Also, it is important that navigation work as expected or a user will become frustrated and lose interest in the application.

@@

\subsection{Selenium\cite{Selenium}}
Selenium is an automated browser input runner. It is one of the only automated browser runners and certainly the most prominent. A developer can write code that will cause Selenium to drive a browser's input and assert expectations. On its own Selenium has very minimal testing usage. It lacks a test reporting mechanism and only offers output in two forms: 1) No output at all, a browser window will open, input will be generated, and the window will close, or 2) the browser will exit with a cryptic error printed to the console. It is up to the developer to include print statements as to what is happening at any given time, what test is being run, and what the results are.

@@

\subsection{Nightwatchjs\cite{NightwatchJS}}
Nightwatch is a JavaScript testing library that wraps Selenium in an effort to make browser testing more concise and straightforward. Nightwatch uses Node.js' assertion library paired with Selenium's WebDriver to both run browser automation and also report test results.

@@

\section{Testing Frameworks}
